### 2015-08-13 doc/thesis/glossary.tex
Fixing some typos and some glossary definitions

\item[contravariant] A part of the type constructor is contravariant when
it reverses the subtyping order.

\item[covariant] A part of the type constructor is covariant when
it preserves the subtyping order.

\item[depth subtyping] An operation over records that allows variance in the type of
record fields.

@@

\item[fixed table type] A table type which guarantees that there are no
keys with a type that is not one of its key types, and that can have
any number of \emph{fixed} or \emph{closed} references point to it.
See complete definition in page \pageref{def:tabletype}.

@@

\item[gradual type system] A type system that uses the consistency relation
instead of type equality to perform static type checking.
See complete definition in page \pageref{sec:gradual}.

@@

\item[invariant] A part of the type constructor is invariant when it forbids variance.
It is also a way to define type equality through subtyping.

@@

\item[open table type] A table type which guarantees that there are no
keys with a type that is not one of its key types, and that only have
\emph{closed} references point to it.
See complete definition in page \pageref{def:tabletype}.

@@

\item[unique table type] A table type which guarantees that there are no
keys with a type that is not one of its key types, and that do not have
any references point to it.
See complete definition in page \pageref{def:tabletype}.

@@

\item[variadic function] A function that can receive an arbitrary number of arguments.

@@

\item[width subtyping] An operation over records that allows the subtype to include
fields that do not exist in the supertype.

>>>>

\item[contravariant] A part of a type constructor is contravariant when it
reverses the subtyping order, that is, the part $T_{1}$ of a type being
a subtype of the corresponding part $T_{2}$ of another type implies that
$T_{2} \subtype T_{1}$.

\item[covariant] A part of a type constructor is covariant when it
preserves the subtyping order, that is, the part $T_{1}$ of a type being
a subtype of the corresponding part $T_{2}$ of another type implies that
$T_{1} \subtype T_{2}$.

\item[depth subtyping] A subtyping relation over records that allows variance
in the type of record fields.

@@

\item[fixed table type] A table type which guarantees that there are no
keys with a type that is not one of its key types, and that can have
any number of \emph{fixed} or \emph{closed} references.
See complete definition in page \pageref{def:tabletype}.

@@

\item[gradual type system] A type system that uses either the consistency relation
or the consistent-subtyping relation instead of type equality to perform
static type checking.
See complete definition in page \pageref{sec:gradual}.

@@

\item[invariant] A part of a type constructor is invariant when it forbids variance,
that is, the part $T_{1}$ of a type being a subtype of the corresponding part $T_{2}$ of
another type implies that $T_{1} \subtype T_{2}$ and $T_{2} \subtype T_{1}$.
It is also a way to define type equality through subtyping.

@@

\item[open table type] A table type which guarantees that there are no
keys with a type that is not one of its key types, and that only have
\emph{closed} references.
See complete definition in page \pageref{def:tabletype}.

@@

\item[unique table type] A table type which guarantees that there are no
keys with a type that is not one of its key types, and that has only one reference.
See complete definition in page \pageref{def:tabletype}.

@@

\item[variadic function] A function that can use an arbitrary number of arguments.

@@

\item[width subtyping] A subtyping relation over records that allows the subtype to
include fields that do not exist in the supertype.

### 2015-08-13 doc/thesis/typedlua.tex
Fixing some typos and some glossary definitions

Typed Lua provides two operation modes: the \emph{default} mode and
the \emph{strict} mode.
In the default mode, the compiler adds \texttt{value*} to the type
of the parameter list and \texttt{nil*} to the return type when
the programmer does not specify a variadic tail. 
It has this behavior to preserve the semantics
of function calls in Lua, that is, it uses \texttt{value*} to discard
extra arguments and \texttt{nil*} to replace missing arguments.
In the strict mode, the compiler adds \texttt{nil*} to both
parts of the function type.
Even though it still uses \texttt{nil*} to replace missing values,
it also uses \texttt{nil*} instead of \texttt{value*} to catch arity mismatch.

We will use the following function to illustrate how these two
operation modes work:
\begin{verbatim}
    local function sum (x:integer, y:integer):integer
      return x + y
    end
\end{verbatim}

@@

Even though these operation modes affect function calls, they do not
affect other multiple assignments.
This means that both operation modes discard extra arguments in
multiple assignments and they also use \texttt{nil} to replace missing
arguments in multiple assignments.
We made this design decision to avoid being to restrictive, that is,
we did not want to give unnecessary warnings to programmers,
as discarded values do not have any implications during run-time.

As an example,
\begin{verbatim}
    local x:integer, y:integer = 1, 2, 3
\end{verbatim}
compiles without any compile-time warnings in both operation modes.
This means that the left-hand side of this local declaration has the
tuple type \texttt{(integer, integer, value*)}, while the expression
list in the right-hand side of this local declaration has the tuple
type \texttt{(1, 2, 3, nil*)} in both operation modes.
The literal type \texttt{3} is consistent with \texttt{value},
and the type \texttt{nil*} generates a value of type \texttt{nil}
that is also consistent with \texttt{value}.

As another example,
\begin{verbatim}
    local x:integer, y:integer = sum(2, 2)
\end{verbatim}
compiles with one compile-time warning in both of the operation modes.
This means that the left-hand side of this local declaration has
the tuple type \texttt{(integer, integer, value*)}, while the
expression list in the right-hand side of this local declaration
has the tuple type \texttt{(integer, nil*)} in both operation modes.
The type \texttt{nil*} generates a value of type \texttt{nil}
that is not consistent with \texttt{integer}, the type of \texttt{y}.

@@

As an example,
\begin{verbatim}
    local x:integer, y:integer, z:string = m(), m()
\end{verbatim}
compiles without compile-time warnings in both operation modes.
This happens because in the right-hand side of the multiple
assignment, only the first value produced by the first call
to \texttt{m} gets used, so the type of the right-hand side
is \texttt{(integer, integer, string, nil*)}, which is
consistent with the type of the left-hand side
\texttt{(integer, integer, string, value*)}.

@@

We can also type variadic functions.
For instance,
\begin{verbatim}
    local function v (...:string):(string*)
      return ...
    end
\end{verbatim}
has type \texttt{(string*) -> (string*)} in both operation modes.
The function call \texttt{v()} compiles without any compile-time warnings in
both modes because the type of the argument list is \texttt{(nil*)},
which is consistent with \texttt{(string*)}.
The call \texttt{v("hello", "world")} also compiles without any warnings
in both modes because the type of the argument list is \texttt{("hello", "world", nil*)},
which is consistent with \texttt{(string*)}.
Calling \texttt{v(...)} compiles without any warnings in both
operation modes because the type of the argument list is \texttt{(string*)},
assuming that the vararg expression (\texttt{...}) has type \texttt{string}.

@@

After we define the function \texttt{message}, we can call it
with a missing argument in both operation modes:
\begin{verbatim}
    print(message("Typed Lua"))         --> Hello Typed Lua
    print(message("Typed Lua", "Hi "))  --> Hi Typed Lua
\end{verbatim}

>>>>

Typed Lua provides two modes of operation: the \emph{default} mode and
the \emph{strict} mode.
In the default mode, the compiler adds \texttt{value*} to the type
of the parameter list and \texttt{nil*} to the return type when
the programmer does not specify a variadic tail. 
It has this behavior to preserve the semantics
of function calls in Lua, that is, it uses \texttt{value*} to discard
extra arguments and \texttt{nil*} to replace missing arguments.
In the strict mode, the compiler adds \texttt{nil*} to both
parts of the function type.
Even though it still uses \texttt{nil*} to replace missing values,
it also uses \texttt{nil*} instead of \texttt{value*} to catch arity mismatch.

We will use the following function to illustrate how these two
modes work:
\begin{verbatim}
    local function sum (x:integer, y:integer):integer
      return x + y
    end
\end{verbatim}

@@

Even though these modes of operation affect function calls, they do not
affect other multiple assignments.
This means that both modes discard extra arguments in
multiple assignments and they also use \texttt{nil} to replace missing
arguments in multiple assignments.
We made this design decision to avoid being too restrictive, that is,
we did not want to give unnecessary warnings to programmers,
as discarded values do not have any implications during run-time.

As an example,
\begin{verbatim}
    local x:integer, y:integer = 1, 2, 3
\end{verbatim}
compiles without any compile-time warnings in both modes of operation.
This means that the left-hand side of this local declaration has the
tuple type \texttt{(integer, integer, value*)}, while the expression
list in the right-hand side of this local declaration has the tuple
type \texttt{(1, 2, 3, nil*)} in both modes.
The literal type \texttt{3} is consistent with \texttt{value},
and the type \texttt{nil*} generates a value of type \texttt{nil}
that is also consistent with \texttt{value}.

As another example,
\begin{verbatim}
    local x:integer, y:integer = sum(2, 2)
\end{verbatim}
compiles with one compile-time warning in both modes of operation.
This means that the left-hand side of this local declaration has
the tuple type \texttt{(integer, integer, value*)}, while the
expression list in the right-hand side of this local declaration
has the tuple type \texttt{(integer, nil*)} in both modes.
The type \texttt{nil*} generates a value of type \texttt{nil}
that is not consistent with \texttt{integer}, the type of \texttt{y}.

@@

As an example,
\begin{verbatim}
    local x:integer, y:integer, z:string = m(), m()
\end{verbatim}
compiles without compile-time warnings in both modes of operation.
This happens because in the right-hand side of the multiple
assignment, only the first value produced by the first call
to \texttt{m} gets used, so the type of the right-hand side
is \texttt{(integer, integer, string, nil*)}, which is
consistent with the type of the left-hand side
\texttt{(integer, integer, string, value*)}.

@@

We can also type variadic functions.
For instance,
\begin{verbatim}
    local function v (...:string):(string*)
      return ...
    end
\end{verbatim}
has type \texttt{(string*) -> (string*)} in both modes of operation.
The function call \texttt{v()} compiles without any compile-time warnings in
both modes because the type of the argument list is \texttt{(nil*)},
which is consistent with \texttt{(string*)}.
The call \texttt{v("hello", "world")} also compiles without any warnings
in both modes because the type of the argument list is \texttt{("hello", "world", nil*)},
which is consistent with \texttt{(string*)}.
Calling \texttt{v(...)} compiles without any warnings in both
operation modes because the type of the argument list is \texttt{(string*)},
assuming that the vararg expression (\texttt{...}) has type \texttt{string}.

@@

After we define the function \texttt{message}, we can call it
with a missing argument in both modes of operation:
\begin{verbatim}
    print(message("Typed Lua"))         --> Hello Typed Lua
    print(message("Typed Lua", "Hi "))  --> Hi Typed Lua
\end{verbatim}
